### 2022-05-09 manuscript.Rmd
Update methods

documentclass     : "apa7"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"
output            : papaja::apa6_pdf
header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage{makecell}
  - \usepackage{setspace}
  - \renewcommand{\arraystretch}{0.5}
  - \DeclareCaptionLabelSeparator{bar}{ | }
  - \DeclareCaptionTextFormat{tabletext}{\textup{#1}}
  - \captionsetup[table]{font=small,labelfont=bf,labelsep=bar,textfont=bf,textformat=tabletext}
  - \captionsetup[figure]{belowskip=10pt,font=small,labelfont=bf,labelsep=bar,name=Fig.,textfont=normalfont}
---

@@

Articles needed to fulfill six criteria for being included in this meta-analysis: (1) The article was written in English or German; (2) The article includes results from a group study with human samples (thus excluding review articles, meta-analyses, case studies, and animal studies); (3) These samples include at least one group of infants (mean age between 0 months and 36 months); (4) Infants were not born preterm and had no clinical diagnosis; (5) Infants performed a mental rotation task; (6) The article contains quantitative scores that could be converted into a standardized mean difference (see Data collection and process items below). We explicitly included works that were not peer reviewed (e.g., dissertations and preprints) to reduce the impact of publication bias.

@@

Many of these articles consisted of multiple experiments, e.g., using different variations of the mental rotation task or different subsamples of infants. We included all of these experiments in the meta-analysis and accounted for the dependencies between them by means of multilevel modeling with by-article random effects (see Bayesian meta-analysis below). However, we excluded experiments if there was insufficient information to compute a standardized effect size (see Data collection process and items below). Whenever an article reported separate effect sizes for males and females---or other subgroups like crawling and non-crawling infants---but also an effect size combining these groups, we only included the combined effect size. Moreover, we disregarded effect sizes that were clearly based on the same data but reported in different articles. This procedure led to a total of `r descs$n_experiments` experiments being included in the meta-analysis (Supplementary Table 1).

@@

Outcome measures and other relevant variables were extracted from each article by one of three raters and verified by a second rater. Outcome measures were any summary statistic (Table 1) that could be used to determine the standardized mean difference between novel/unexpected rotation events and familiar/expected rotation events (see Introduction). Other extracted variables included, if available, the sample size, the number of males and females, the mean age and its standard deviation, the minimum and maximum age, the type of mental rotation task (habituation or violation of expectation), the modality of stimulus presentation (real objects or objects on a computer screen), and the dimensionality of the stimuli (2D or 3D; Supplementary Table 1).

@@

One outcome measure per experiment was converted into a standardized mean difference with small sample correction (Hedges’ $g$) using the formulas provided in Table 1. The standard error $SE_g$ of the outcome measure $g$ for each experiment was computed using the formula provided by @goulet-pelletier2018:

$${SE}_\text{rotation}=\sqrt{\frac{{df}}{{df}-2}\frac{2(1-r)}{n}\left(1+g_\text{rotation}^2\frac{n}{2(1-r)}\right)-\frac{g_\text{rotation}^2}{J^2}}$$

where $n$ is the sample size of the experiment, ${df}$ are the degrees of freedom (with ${df} = 2*(n-1)$), $r$ is the correlation between the two dependent measures in the experiment, and $J({df})$ is the correction factor for small samples as described in Table 1. The correlation $r$ was not reported in any of the original articles [@harrer2021]. We therefore always assumed a correlation of $r = .50$ to make our analysis comparable to standard (between-group) meta-analyses [@cohen1988; @lakens2013; @morris2002] and because we were able to infer an average correlation of $r \approx .50$ from a subsample of articles which provided sufficient information (Supplementary Methods 1). A sensitivity analysis indicated that changing the assumed correlation to values from $r = -.90$ via $r = .00$ to $r = .90$ had no meaningful impact on the meta-analytic effect size (Supplementary Table 5).

For the meta-analysis of gender differences within each article, effect sizes were computed according to the following formulas in Table yy, and redundant groups were excluded from the analysis.

@@

Furthermore, the sampling variance was computed according to the following equation:

$${SE}_\text{gender}=\sqrt{\frac{{df}}{{df}-2}\frac{2}{\tilde{n}}\left(1+g_\text{gender}^2\frac{\tilde{n}}{2}\right)-\frac{g_\text{gender}^2}{J({df})^2}}$$

Where $\tilde{n}$ is the harmonic mean of the group sizes
(i.e., $\tilde{n}=\frac{2}{n_\text{female}^{-1}+n_\text{male}^{-1}}$)
and ${df}$ are the degrees of freedom as provided in row (1) of Table 2.
A sensitivity analysis highlights the stability of the effect size computed using the Bayesian meta analysis (Supplementary Table 5).

@@

We synthesized the effect sizes and their sampling variances using a Bayesian multilevel model. This model had three levels, with infant participants nested in experiments and experiments nested in articles [@vandennoortgate2013]. We used a weakly-informative $\mathcal{N}(0,1)$ (normal) prior for the meta-analytic effect size and a weakly-informative $\mathcal{HC}(0,0.3)$ (half-Cauchy) prior for all standard deviations [@williams2018]. A prior sensitivity analysis indicated that making these priors either more informative or less informative did not change the meta-analytic results (Supplementary Table 5). All Bayesian models were fitted using the brms package (Version `r packageVersion("brms")`) [@burkner2017; @burkner2018] in R (Version `r packageVersion("base")`) [@rcoreteam2021] and the Stan language (Version `r packageVersion("rstan")`) [@standevelopmentteam2022]. Markov Chain Monte Carlo (MCMC) sampling was used with four parallel chains, each sampling `r formatC(n_iter, big.mark = ",")` draws (including `r formatC(n_warmup, big.mark = ",")` warm-up draws) from the posterior distribution. To verify the convergence of the Markov chains, we examined rank plots as well as the $R_\text{hat}$ and $N_\text{eff}$ statistics (Supplementary Fig. 2) [@vehtari2021]. For reporting, the credible interval (CrI) for each model parameter was computed as the 95% equal-tailed interval (ETI) of its posterior distribution, although replacing this with the 95% highest density interval (HDI) yielded highly similar results [@kruschke2015].

@@

Publication bias was evaluated based on a funnel plot and an Egger regression test [@egger1997; @sterne2005]. To produce the funnel plot, which visualizes the relationship between standard errors and effect sizes, we adapted code from the R package metaviz (`r packageVersion("metaviz")`) [@kossmeier2020]. For the Egger regression test we used the R package metafor (Version `r packageVersion("metafor")`) [@viechtbauer2010] to identify a potential relationship between standard errors and effect sizes. Such a relationship is indicated by the observation that the slope of a weighted linear regression of the effect sizes on the standard errors is significantly different from zero in a two-sided $t$-test.

To scrutinize the robustness of the meta-analytic effect size against the effects of individual studies we conducted a jackknife leave-one-out-analysis [@gee2005; @efron1993]. To this end, we adapted code from the bootstrap R package (Version `r packageVersion("bootstrap")`) [@kostyshak2019]. Specifically, we fit the Bayesian models repeatedly to our study samples, leaving out one experiment sample for every iteration. To take into account the variance inherent in Bayesian model estimation, we averaged over ??? iterations of the jackknife.

>>>>

documentclass     : "apa7"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"
output            : papaja::apa6_pdf
header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage{makecell}
  - \usepackage{setspace}
  - \renewcommand{\arraystretch}{0.5}
  - \DeclareCaptionLabelSeparator{bar}{ | }
  - \DeclareCaptionTextFormat{tabletext}{\textup{#1}}
  - \captionsetup[table]{belowskip=20pt,font=small,labelfont=bf,labelsep=bar,textfont=bf,textformat=tabletext}
  - \captionsetup[figure]{belowskip=10pt,font=small,labelfont=bf,labelsep=bar,name=Fig.,textfont=normalfont}
---

@@

Articles needed to fulfill six criteria for being included in this meta-analysis: (1) The article was written in English or German; (2) The article includes results from a group study with human samples (thus excluding review articles, meta-analyses, case studies, and animal studies); (3) These samples include at least one group of infants (mean age between 0 months and 36 months); (4) Infants were not born preterm and had no clinical diagnosis; (5) Infants performed a mental rotation task; (6) The article contains quantitative scores that could be converted into a standardized mean difference (see Data collection process and items below). We explicitly included works that were not peer reviewed (e.g., dissertations and preprints) to reduce the impact of publication bias.

@@

Many of these articles consisted of multiple experiments, e.g., using different variations of the mental rotation task or different subsamples of infants. We included all of these experiments in the meta-analysis and accounted for the dependencies between them by means of multilevel modeling with by-article random effects (see Bayesian meta-analysis below). However, we excluded experiments if there was insufficient information to compute a standardized effect size (see Data collection process and items below). Whenever an article reported separate effect sizes for males and females---or other subgroups like crawling and non-crawling infants---but also an effect size combining these groups, we only included the combined effect size. Moreover, we disregarded effect sizes that were clearly based on the same data but reported in different articles. This procedure led to a total of `r descs$n_experiments` experiments being included in the meta-analysis of mental rotation performance (Supplementary Table 1) and 30 articles being included in the meta-analysis of gender differences in mental rotation performance.

@@

Outcome measures and other relevant variables were extracted from each article by one of three raters and verified by a second rater. For the meta-analysis of mental rotation performance, outcome measures were any summary statistic (Table 1) that could be used to determine the standardized mean difference between novel/unexpected rotation events and familiar/expected rotation events (see Introduction). Other extracted variables included, if available, the sample size, the number of males and females, the mean age and its standard deviation, the minimum and maximum age, the type of mental rotation task (habituation or violation of expectation), the modality of stimulus presentation (real objects or objects on a computer screen), and the dimensionality of the stimuli (2D or 3D; Supplementary Table 1).

@@

We also conducted a meta-analysis of the gender differences in mental rotation performance observed within the original articles. For this analysis, the outcome measures were any summary statistic (Table 2) that could be used to determine the standardized mean difference between male infants’ mental rotation performance and female infants’ mental rotation performance. Other extracted variables included the sample size, the mean age and its standard deviation, and the minimum and maximum age of each gender group.

@@

For the meta-analysis of mental rotation performance, one outcome measure per experiment was converted into a standardized mean difference with small sample correction (Hedges’ $g$) using the formulas provided in Table 1. The standard error of Hedges' $g$ for each experiment was computed using the formula provided by @goulet-pelletier2018: $${SE}_\text{rotation}=\sqrt{\frac{{df}}{{df}-2}\frac{2(1-r)}{n}\left(1+g_\text{rotation}^2\frac{n}{2(1-r)}\right)-\frac{g_\text{rotation}^2}{J^2}}$$ where $n$ is the sample size of the experiment, ${df}$ are the degrees of freedom (with ${df} = 2*(n-1)$), $r$ is the correlation between the two dependent measures in the experiment, and $J({df})$ is the correction factor for small samples as described in Table 1. The correlation $r$ was not reported in any of the original articles [@harrer2021]. We therefore always assumed a correlation of $r = .50$ to make our analysis comparable to standard (between-group) meta-analyses [@cohen1988; @lakens2013; @morris2002] and because we were able to infer an average correlation of $r \approx .50$ from a subsample of articles which provided sufficient information (Supplementary Methods 1). A sensitivity analysis indicated that changing the assumed correlation to values from $r = -.90$ via $r = .00$ to $r = .90$ had no meaningful impact on the meta-analytic effect size (Supplementary Table 5).

For the meta-analysis of gender differences in mental rotation performance, one outcome measure per contrast between non-redundant male and female groups of infants was converted into a standardized mean difference with small sample correction (Hedges’ $g$) using the formulas provided in Table 2. The standard error of Hedges’ $g$ for each gender contrast was computed using the formula provided by @goulet-pelletier2018: $${SE}_\text{gender}=\sqrt{\frac{{df}}{{df}-2}\frac{2}{\tilde{n}}\left(1+g_\text{gender}^2\frac{\tilde{n}}{2}\right)-\frac{g_\text{gender}^2}{J({df})^2}}$$ where $df$ are the degrees of freedom (with $df=n_\text{female}+n_\text{male}-2$), $\tilde{n}$ is the harmonic mean of the group sizes (i.e., $\tilde{n}=\frac{2}{n_\text{female}^{-1}+n_\text{male}^{-1}}$), and $J(df)$ is the correction factor for small samples as described in Row (1) of Table 2.

@@

We synthesized the effect sizes and their sampling variances using a Bayesian multilevel model. This model had three levels, with infant participants nested in experiments and experiments nested in articles [@vandennoortgate2013]. We used a weakly-informative $\mathcal{N}(0,1)$ (normal) prior for the meta-analytic effect size and a weakly-informative $\mathcal{HC}(0,0.3)$ (half-Cauchy) prior for all standard deviations [@williams2018]. A prior sensitivity analysis indicated that making these priors either more informative or less informative did not change the meta-analytic results (Supplementary Table 5). For the meta-analysis of mental rotation performance, the dependent variable was the standardized mean difference (Hedges’ g) between the novel and familiar rotation condition, weighted by its standard error (see Effect size measures above). For the meta-analysis of gender differences, the dependent variable was the standardized mean difference (Hedges’ g) between male infants’ mental rotation performance and female infants’ mental rotation performance, weighted by its standard error. All Bayesian models were fitted using the brms package (Version `r packageVersion("brms")`) [@burkner2017; @burkner2018] in R (Version `r packageVersion("base")`) [@rcoreteam2021] and the Stan language (Version `r packageVersion("rstan")`) [@standevelopmentteam2022]. Markov Chain Monte Carlo (MCMC) sampling was used with four parallel chains, each sampling `r formatC(n_iter, big.mark = ",")` draws (including `r formatC(n_warmup, big.mark = ",")` warm-up draws) from the posterior distribution. To verify the convergence of the Markov chains, we examined rank plots as well as the $R_\text{hat}$ and $N_\text{eff}$ statistics (Supplementary Fig. 2) [@vehtari2021]. For reporting, the credible interval (CrI) for each model parameter was computed as the 95% equal-tailed interval (ETI) of its posterior distribution, although replacing this with the 95% highest density interval (HDI) yielded highly similar results [@kruschke2015].

@@

Publication bias was evaluated based on funnel plots and Egger regression tests [@egger1997; @sterne2005]. Funnel plots visualize the relationship between standard errors and effect sizes. They were created by adapting code from the R package metaviz (`r packageVersion("metaviz")`) [@kossmeier2020]. The Egger regression test is a formal statistical test for this relationship between standard errors and effect sizes, probing if the weighted linear regression weight of the effect sizes on the standard errors is significantly different from zero. This test was performed using code from the metafor package and applying a two-sided false-positive error rate of $\alpha = .05$.

To scrutinize the robustness of the meta-analytic effect size against the influence of any individual experiment (which may or may not be a false-positive), we conducted a jackknife leave-one-out-analysis for the meta-analysis of mental rotation performance [@gee2005; @efron1993]. To this end, we adapted code from the bootstrap R package (Version `r packageVersion("bootstrap")`) [@kostyshak2019]. Specifically, we refitted the Bayesian three-level models repeatedly while leaving out a different one of the original experiments on every iteration. To take into account the variance inherent in Bayesian model estimation, we averaged over $x$ iterations of the jackknife.
